Naturalness in U(1) extensions of the MSSM

### 2015-01-20 paper/e6ssm-tuning-paper.tex
Edited introduction for clarity and to account for what we now have.
E6SSM parts still need to be heavily edited.

The Higgs causes a naturalness problem because at tree level it has an
upper bound of the mass of the $Z$ boson, $M_Z$. The dominant higher
order corrections to the Higgs mass come from stops and to obtain a
$125$ GeV Higgs they need be rather heavy. Heavy stops will provide a
large contribution to the soft breaking mass for the up-type Higgs
scalar, $m_{H_u}^2$ at low energies, through the evolution of the
renormalisation group equations (RGEs) from the grand unification
(GUT) scale to the electroweak (EW) scale. This affects the SUSY
prediction of the electroweak VEV, $v$, or $M_Z$.  This naturalness
problem motivates both further examination of non-minimal SUSY models
that can raise the Higgs mass without the need for heavy stops and
alternative possibilities for how the soft breaking parameters get
generated, which might set them them at lower energies, reducing the
influence the stops have on the soft breaking mass for the up-type
Higgs scalar, $m_{H_u}^2$.

@@

In this paper we investigate how big this tuning is if we bring this
scale all the way down to 20 TeV, effectively minimising contribution
from the stops. We find that the $Z^\prime$ limit is enough to already
require moderate fine tuning.  We also show this is comparable to the
situation in the MSSM defined at the same scale if chargino limit is
$700$ GeV, which is the most conservative limit on can obtain and
above the tuning from the stops, in both the MSSM and U(1) extensions.
We also show that the tuning from the stops becomes larger when the
high scale is raised to ?? whereas for the E$_6$SSM the minimum tuning
is lower than the MSSM above this scale.  Therefore it is clear that
which model is preferred and how large the tuning is depends on our
assumptions about SUSY breaking models.

However we also show that if the $Z^\prime$ limit is the extended to
?? then $U(1)$ extensions of the MSSM will already suffer from a
severe tuning, irrespective of the UV boundary condition for the RGE
evolution.  Therefore the $Z^\prime$ constraint is amongst the most
important in terms of tuning and attacking natural supersymmetry
experimentally and the next run of the LHC will be crucial.

Here we will refer to U(1) extensions of the MSSM, which solve the
$\mu$-problem as the USSM \cite{Cvetic:1995rj, Jain:1995cb,
  Nir:1995bu, Cvetic:1996mf, Cvetic:1997ky}, while we will sometimes
specify that the models are inspired by $E_6$ by referring to them as
such. Finally we will also discuss specific models with complete $E_6$
matter multiplets at low energies and an extra $U(1)$ under which the
right handed neutrino remains interaction-less as
E$_6$SSM-variants\cite{Howl:2007zi, Braam:2009fi, Braam:2010sy,
  Hall:2011zq, Nevzorov:2012hs, Athron:2014pua}, with the E$_6$SSM
reserved for the original formulation of this model
\cite{King:2005jy,King:2005my,Athron:2010zz} original formulation of
the model.

>>>>

In the MSSM the Higgs mass causes a naturalness problem because at
tree level it has an upper bound of the mass of the $Z$ boson,
$M_Z$. The dominant higher order corrections to the Higgs mass come
from stops and to obtain a $125$ GeV Higgs they need be rather
heavy. Heavy stops will provide a large contribution to the low energy
value of $m_{H_u}^2$, the soft breaking mass for the up-type Higgs
scalar, through the evolution of the renormalisation group equations
(RGEs) from the grand unification (GUT) scale to the electroweak (EW)
scale. This affects the SUSY prediction of the electroweak VEV, $v$,
or $M_Z$.  This naturalness problem motivates both further examination
of non-minimal SUSY models that can raise the Higgs mass without the
need for heavy stops and alternative possibilities for how the soft
breaking parameters get generated, which might set them at lower
energies, reducing the influence the stops have on $m_{H_u}^2$.

@@

In this paper we investigate how big this tuning is if we bring this
scale all the way down to 20 TeV, effectively minimising contribution
from the stops. We find that the $Z^\prime$ limit is enough to already
require moderate fine tuning in the E6SSM.  We also show this is
comparable to the situation in the MSSM defined at the same scale if
chargino's could be ruled out below $700$ GeV.  We then show how this
tuning from the $Z^\prime$ mass looks for different $U(1)$ extensions,
finding that the current severity depends upon the charges but that
$Z^\prime$ limits are important in constraining the most natural
scenarios in all of these models.  Therefore the $Z^\prime$ constraint
is amongst the most important in terms of tuning and attacking natural
supersymmetry experimentally and the next run of the LHC will be
crucial in this respect.

Finally we make a case study, for a few benchmarks, of the impact of
raising the high scale boundary condition, $M_X$, at which the SUSY
breaking parameters must be fixed by some SUSY breaking mechanism. We
show that which model has less fine tuning depends on $M_X$.  We also
see rather complicated behaviour in the tuning for the E$_6$SSM points
due to the combination of different sources of tuning.

Here we will refer to U(1) extensions of the MSSM, which solve the
$\mu$-problem as the USSM \cite{Cvetic:1995rj, Jain:1995cb,
  Nir:1995bu, Cvetic:1996mf, Cvetic:1997ky}, but since avoiding gauge
anomalies is straightforward in the when the $U(1)$appears from the
breakdown of $E_6$ and complete matter multiplets are present at low
energies, we will sometimes refer to such models as $E_6$ inspired.
Models with with complete $E_6$ matter multiplets at low energies and
an extra $U(1)$ under which the right handed neutrino remains
interaction-less are often refered to as the exceptional supersymmtric
standard models (E$_6$SSM). However we will reserve the name E$_6$SSM
to the original formulation of models
\cite{King:2005jy,King:2005my,Athron:2010zz}, and refer to similar
models as E$_6$SSM-variants\cite{Howl:2007zi, Braam:2009fi,
  Braam:2010sy, Hall:2011zq, Nevzorov:2012hs, Athron:2014pua}.
